### 2018-02-14 Writing/Thesis/content/abstract.tex
more writing
more table and figure overhaul
First draft

Should abstract be included?

\vspace*{20mm}

{\usekomafont{chapter}Abstract (different language)}\label{sec:abstract-diff} \\

Und falls ja, dann zweisprachig?

>>>>

Although the amount of data available worldwide is increasing, the portion of analyzed data remains low. This problem is caused by a lack of experts in the field that are able to carry out an analysis, and the difficulty of the analysis for non-experts. Especially for the task of classification, there are many algorithms with variing performance values from data set to data set - there does not exist one best algorithm, which one can always recommend. One response to this problem is automated machine learning. Therefore, in this thesis, to approaches to construct a ranking of classification algorithms for a new data set, based on properties of that data set,are presented and compared. The first of the two alternatives works on the basis of regression models; one regression model predicts a performance value for each classifier, from which then a ranking is constructed. In contrast, the second, preference based model directly learns and returns rankings. In the evaluation it became clear that the regression based alternative is superior to the preference based approach; the best evaluated regression model, based on the random forest algorithm, achieved values of up to 0.495 for Kendall's rank correlation coefficient and a peformance loss of 1.308 percent on average. However, the tested preference models are superior at least to the baseline in some scenarios. Therefore the conlucion can be drawn that at least certain properties of data sets are connected to performance values of classification algorithms, which can be exploited by regression and preference models, although the regression approach is superior to the preference alternative within the scope of the conducted evaluation.

\vspace*{20mm}
\newpage

{\usekomafont{chapter}Abstract (German)}\label{sec:abstract-german} \\

% Motivation
Die weltweit verfügbare Menge an Daten steigt immer weiter an, und doch bleibt der Anteil der tatsächlich analysierten Daten gering. Dieses Problem entsteht, da nicht genug Experten verfügbar sind, um diese Analyse auszuführen, und es für Laien schwierig ist, das richtige Verfahren auszuwählen. Besonders im Bereich der Klassifikation gibt es viele Algorithmen, deren Performanz jedoch von Datensatz zu Datensatz variieren kann - es gibt also keinen einen besten Algorithmus, der immer eine gute Wahl darstellt. Eine Antwort auf diese Problem ist automatisiertes Machine Learning. Daher werden in dieser Arbeit zwei verschiedene Ansätze, welche für einen neuen Datensatz ein Ranking von Klassifizierungsalgorithmen aufgrund von Eigenschaften des Datensatzes vorhersagt, vorgestellt und miteinander verglichen. Der erste der beiden Ansätze arbeitet auf Grundlage von Regressionsmodellen; ein Ranking wird hier konstruiert indem ein einzelnes Regressionsmodell für jeden Klassifizierungsalgorithmus einen Performanzwert vorhersagt, auf deren basierend dann das Ranking erstellt wird. Der zweite Ansatz beschäftigt sich mit Präferenzmodellen, welche direkt Rankings lernen und zurückgeben. In der Evaluation stellte sich heraus, dass der Regressionsansatz dem Präferenzmodell überlegen ist; das beste der getesteten Regressionsmodelle, basierend auf dem Random Forest Algorithmus, erreicht einen Wert von Kendalls Tau von bis zu 0,495 und Performanzverlustwerte von durchschnittlich nur 1,308 Prozentpunkten. Jedoch sind auch manche der getesten Präferenzmodelle in einigen Punkten wenigstens der zum Vergleich benutzten Baseline überlegen. Daher kann die Schlussfolgerung gezogen werden, dass wenigstens einige der benutzen Eigenschaften der Datensätze mit den Performanzwerten der Klassifizierungsalgorithmen zusammenhängen, und dieser Zusammenhang durch Regressionsmodelle sowohl als auch Präferenzmodelle ausgenutzt werden kann, wobei jedoch der Regressionsansatz letzterem im Rahmen der durchgeführten Evaluation überlegen ist.

### 2018-02-14 Writing/Thesis/content/chapter-approach.tex
First draft

In this chapter the two alternatives to tackle the problem of ranking classification algorithms regarding properties of the data set based on past performance values are described in detail. First, the theoretical approach is presented. Then, implementation details are discussed.

@@

Approach of ranking classifier implementations here realized by two concepts, regression-based and preference-based ranking, which will be explained in detail. Also predicting predictive accuracy here. Thus target function of the form $meta features^n \rightarrow classifiers^m$.
performance measure chosen predictive accuracy

% Theory of approach
Since the aim is to generate a ranking of classification algorithms, first, performance values of a number of classifiers are recorded on a few data sets. Then, meta features are computed for each data set. This is for training the ranking model. This generates a table that contains all necessary informatino for training both types of rankers. A ranking of the classifiers for a new instance is then achieved by computing the meta features for the data set and subsequently using it to query the underlying model.

@@

\subsection{Regression Based Ranking}
% How is the regression ranking done
One possibility to derive a ranking of classifiers from this information when given a new data set is to use regression models. The idea is to use separate regression models to predict a performance value for each classifier, and to then derive an ordering from these predictions. This is done by splitting the data set compiled beforehand into separate data sets that each contain all meta features for all data sets, but only the performance values of one classifier. The target feature on these individual data sets, the performance value of the classifier, is a numeric value, and thus a regression model can be trained on each of them. All regression models therefore try to learn the target function $meta features \rightarrow classifier performance value$ for their respective classifier. Hence for a query instance, after computing the meta features, these are fed into each regression model and the predicted performance value for each classifier is saved. In a second step, the classifiers are ordered according to the predictions.

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
	\begin{tabularx}{1.1\textwidth}{X | X | X | X}
		%\hline
		Meta Feature 1	& ... 	& Meta Feature n		& Performance Value 1	\\ \hline
		$v_{11}$			& ...	& $v_{1n}$			& $p_{11}$				\\ 
		$v_{21}$			& ...	& $v_{2n}$			& $p_{21}$				\\ 
		...				& ...	& ...				& ...					\\ 
		$v_{m1}$			& ... 	& $v_{mn}$			& $p_{m1}$				\\ \hline\hline
		Meta Feature 1	& ... 	& Meta Feature n		& Performance Value 2 	\\ \hline
		$v_{11}$			& ...	& $v_{1n}$			& $p_{12}$				\\ 
		$v_{21}$			& ...	& $v_{2n}$			& $p_{22}$				\\ 
		...				& ...	& ...				& ...					\\ 
		$v_{m1}$			& ... 	& $v_{mn}$			& $p_{m2}$					 
	\end{tabularx}
}	
\vspace{1em}
...
\vspace{1em}	
\resizebox{\textwidth}{!}{%
	\begin{tabularx}{1.1\textwidth}{X | X | X | X}
		%\hline
		Meta Feature 1	& ... 	& Meta Feature n		& Performance Value k 	\\ \hline
		$v_{11}$			& ...	& $v_{1n}$			& $p_{1k}$				\\ 
		$v_{21}$			& ...	& $v_{2n}$			& $p_{2k}$				\\ 
		...				& ...	& ...				& ...					\\ 
		$v_{m1}$			& ... 	& $v_{mn}$			& $p_{mk}$					 
	\end{tabularx}
}
	\label{tab:regressionTables}
	\caption{How the regression ranker processes the training data.}
\end{table}

@@

\subsection{Preference Based Ranking}
% How is the preference ranking done
The second possibility considered in this context is using preference learning to predict a ranking. Each instance of the data set generated beforehand contains meta feature information for the considered data set and performance values of all classifiers. Similarly to the regression alternative, the performance values can be converted to preference information by sorting the corresponding classifiers by their performance values. This leads to an ordering of classifiers associated with each instance, and implies that the preference learning task at hand is label ranking. In this case, there exists only one label ranking mode which is fed the computed meta data for a new data set and returns an ordering of classifiers. Thereby it attempts to learn the mapping from meta features to an ordering of classifiers directly.

\begin{table}[h]
\centering
	\begin{tabularx}{\textwidth}{X | X | X | X | X | X | X | X}
		%\hline
		MF 1				& MF 2				& ... 	& MF n				& PV 1 		& PV 2 		&	...	&	PV k 		\\ \hline
		$v_{11}$			& $v_{12}$			& ...	& $v_{1n}$			& $p_{11}$	& $p_{12}$	& 	...	&	$p_{1k}$		\\ 
		$v_{21}$			& $v_{22}$			& ...	& $v_{2n}$			& $p_{21}$	& $p_{22}$	& 	...	&	$p_{2k}$		\\
		...				& ...				& ...	& ...				& ...		& ...		&	...	&	...			\\
		$v_{m1}$			& $v_{m2}$			& ... 	& $v_{mn}$			& $p_{2m}$	& $p_{m2}$	& 	...	&	$p_{mk}$			 
	\end{tabularx}
	\label{tab:preferenceTable}
	\caption{How the preference ranker processes the training data.}
\end{table}

@@

\section{Implementation}
% regression based model - weka used

% preference learning
The jPL framework is a java framework for the evaluation of preference learning algorithms. It implements several tasks from the context of preference learning, including label ranking.

% meta features - openML webapp
\cite{OpenML2013}
% Which meta features
As for the meta features, all meta features discussed in chapter \ref{sec:approach} where used. Initially, it was intended to download the computed meta features from OpenML, but since missing values for meta features ranged from roughly 25\% up to 89\% (except for a few simple meta features which were included for all data sets) it was decided to compute them anew. Meta features where calculated with the help of an implementation provided by OpenML \cite{openMLEvaluationEngine}. Two data sets, one including the downloaded meta features (for all 812 data sets) and one including the computed meta features (for the reduced set of 448 data sets) are thus included in the supplementary material. Since the meta data also includes probing, which is generally an expensive operation and furthermore does not aid in proving the point of a connection of properties of a data set and the performance of classifiers, additionally to the full meta features all evaluation measures where computed without using those meta features that use probing.

>>>>

In this chapter the two alternatives considered in this thesis to tackle the problem of ranking classification algorithms regarding properties of the data set. First, the theoretical approach is presented. Then, implementation details are discussed.

@@

In general, the aim is to predict a ranking of algorithms based on properties, or meta features of a data set. This means, that as training data for a ranking approach (ranker), a number of performance values of classifiers need to be recorded on data sets in a first pre-processing phase, for which also meta features are to be computed. The table of this collected data is illustrated in Table \ref{tab:performanceValues}, for n meta features, k classifiers and n data sets. This approach to ranking is realized here, as pointed out before, in a regression based and preference based approach, which are explained in detail in the following sections.

@@

\subsection{Regression Based Ranking}
% How is the regression ranking done
One possibility to derive a ranking of classifiers from this information when given a new data set is to use regression models. The idea is to use separate regression models to predict a performance value for each classifier, and to then derive an ordering from these predictions. This is done by splitting the data set compiled beforehand (see Tab. \ref{tab:performanceValues}) into separate data sets that each contain all meta features for all data sets, but only the performance values of one classifier, as illustrated in Table \ref{tab:regressionTables}. The target feature on these individual data sets, the performance value of the classifier, is a numeric value, and thus a regression model can be trained on each of them. All regression models therefore try to learn the target function metafeatures$\rightarrow$performance value for their respective classifier. Hence for a query instance, after computing the meta features, these are fed into each regression model and the predicted performance value for each classifier is saved. In a second step, the classifiers are ordered according to the predictions. This process the ranker passes through for making a prediction is presented in Figure \ref{fig:regression_ranker_model}.

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
	\begin{tabularx}{1.1\textwidth}{X | X | X | X}
		%\hline
		Meta Feature 1	& ... 	& Meta Feature n		& Performance Value 1	\\ \hline
		$v_{11}$			& ...	& $v_{1n}$			& $p_{11}$				\\ 
		$v_{21}$			& ...	& $v_{2n}$			& $p_{21}$				\\ 
		...				& ...	& ...				& ...					\\ 
		$v_{m1}$			& ... 	& $v_{mn}$			& $p_{m1}$				\\ \hline\hline
		Meta Feature 1	& ... 	& Meta Feature n		& Performance Value 2 	\\ \hline
		$v_{11}$			& ...	& $v_{1n}$			& $p_{12}$				\\ 
		$v_{21}$			& ...	& $v_{2n}$			& $p_{22}$				\\ 
		...				& ...	& ...				& ...					\\ 
		$v_{m1}$			& ... 	& $v_{mn}$			& $p_{m2}$					 
	\end{tabularx}
}	
\vspace{1em}
...
\vspace{1em}	
\resizebox{\textwidth}{!}{%
	\begin{tabularx}{1.1\textwidth}{X | X | X | X}
		%\hline
		Meta Feature 1	& ... 	& Meta Feature n		& Performance Value k 	\\ \hline
		$v_{11}$			& ...	& $v_{1n}$			& $p_{1k}$				\\ 
		$v_{21}$			& ...	& $v_{2n}$			& $p_{2k}$				\\ 
		...				& ...	& ...				& ...					\\ 
		$v_{m1}$			& ... 	& $v_{mn}$			& $p_{mk}$					 
	\end{tabularx}
}
	\caption{How the regression ranker processes the training data.}
	\label{tab:regressionTables}
\end{table}

@@

\subsection{Preference Based Ranking}
% How is the preference ranking done
The second possibility considered in this context is using preference learning to predict a ranking. Each instance of the data set generated beforehand (see Fig \ref{tab:performanceValues}) contains meta feature information for the considered data set and performance values of all classifiers. Similar to the regression alternative, the performance values can be converted to preference information by sorting the corresponding classifiers by their performance values for each data set. This leads to an ordering of classifiers associated with each instance, and implies that the preference learning task at hand is label ranking. In this case, there exists only one label ranking model which is fed the computed meta data for a new data set and returns an ordering of classifiers, which is depicted in Figure \ref{fig:preference_ranker_model}. Thereby it attempts to learn the mapping from meta features to an ordering of classifiers directly.

\begin{table}[h]
\centering
	\begin{tabularx}{\textwidth}{X | X | X | X | X}
		%\hline
		MF 1				& MF 2				& ... 	& MF n				& Ranking 	\\ \hline
		$v_{11}$			& $v_{12}$			& ...	& $v_{1n}$			& $r_1$		\\ 
		$v_{21}$			& $v_{22}$			& ...	& $v_{2n}$			& $r_2$		\\
		...				& ...				& ...	& ...				& ...		\\
		$v_{m1}$			& $v_{m2}$			& ... 	& $v_{mn}$			& $p_m$		 
	\end{tabularx}
	\label{tab:preferenceTable}
	\caption{How the preference ranker processes the training data.}
\end{table}

@@

\section{Implementation}
In general, all implementations for the approach are carried out in Java, particularly Java 8. However, some additional packages where used, which are pointed out briefly. For the implementation of the regression models, WEKA was used, a popular libary of machine learning algorithms that includes algorithms for classification, regression, and provides help in the evaluation of these learning algorithms \cite{hall2009weka}. Regarding the preference algorithms, the jPL framework was used \cite{intelligent2017jpl}. The jPL framework is a java framework for the evaluation of preference learning algorithms, it implements several tasks from the context of preference learning, including label ranking, and also provides a representation for data sets that contain label ranking information. The list of meta features used is taken from OpenML \cite{OpenML2013}, a website that not only provides a large number of openly accessible data sets, but also records performance values of learning algorithms on these data sets, and features properties of data sets. Thus the meta features where calculated with the help of an implementation provided by OpenML in its Evaluation Engine \cite{openMLEvaluationEngine}; The full list of meta features is included in Table \ref{tab:metaFeatureDetails} in the Appendix.

### 2018-02-14 Writing/Thesis/content/chapter-evaluation.tex
more table and figure overhaul
First draft

% How base performance data generated
The performance data of the classifiers for the selected data sets was generated on a linux cluster, each node consisting of two Intel(R) Xeon(R) E5-2670 processors running at 2.6GHz, 16GB RAM with usage of up to 150 nodes. There was a timeout of eight hours set for the computation of a single performance value of a classifier on a data set. For each of the classifiers, the predictive accuracy (the percentage of correctly classified instances in the data set) was recorded on each data set by means of five times stratified Monte-Carlo cross validation with a 70/30 split in training and test data. When a timeout occurred or the evaluation failed otherwise, the predictive accuracy was set to zero.

%TODO integrate DONT FORGET META FEAUTRES (WHY EVERYTHING X2??)
for mann-whitney U apache commons implementation used with ties averaged and NaNs removed before analysis
apache commons implementation for Kendall Rank correlation as well
JAICore implementation for stratified split of data set

% How rest of the evaluation carried out
All other experiments where carried out on a Windows Machine running Windows 10 Education (version 1709) with a Intel(R) Core(TM) i5-4200U CPU running at 1.60GHz (2.30GHz max) with 12 GB RAM. It is notable that especially the time-related results are to be viewed in regard to this setup. No timeouts where used in this evaluation, and the rankers where evaluated on the training data by means of leave-one-out estimation, that is for the n data sets for which performance values where recorded, they where trained with the meta data - performance examples for n - 1 data sets, and queried for the remaining data set. As n = 448 in this case, the detailed values for each data set are not included in the discussion of the results. They can, however, be found in the supplementary material, offering clues to how each ranker performed in the prediction of a ranking for each of the used data sets, identified by their ID on OpenML.

@@

Compared to the evaluation on the full set of meta data, the results on the reduced set without out probing are less distinctive. For the Kendall Rank correlation, linear regression, M5P and random forest are significantly better than the baseline, and for label ranking, both of the instance based approaches with non-default parameters as well, while the other label ranking approaches are significantly worse than the baseline. In total, random forest is still the best alternative with a correlation of 0.439, slightly worse than the correlation achieved on the full meta data. For both loss measures however, while the disadvantages of the label ranking approaches remain, none of the eight alternatives achieves a significantly better loss or best three loss than the baseline anymore.

Summarizing the observed results, in general, a regression based approach shows an indication to be stronger than a preference based approach. Specifically, using random forest as a regression model is the best solution observed in the evaluation, and label ranking by pairwise comparison the weakest, as it falls behind the other ranking approaches and the baseline often.

@@

\tikzsetnextfilename{LossGraphics}
\begin{figure}
%\pgfplotsset{width=\textwidth}
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}
\begin{axis}[
	title=regular scale,
	height=\textwidth,
	width=\textwidth,
	ylabel={BestThreeLoss}, 
	xlabel={Data Set},
]
\addplot [mark=*,only marks,mark size=1pt] table [x=ID,y=BestThreeLoss,col sep=semicolon] {data/BestAlgorithmRanker_metaData_small_allPerformanceValues.csv};
\addplot [mark=*,only marks,mark size=1pt,uniaccentblue] table [x=ID,y=BestThreeLoss,col sep=semicolon] {data/RandomForestRanker_metaData_small_allPerformanceValues.csv};
\addplot [very thick,mark=none,black,samples=2,domain=0:450]{2.440};
\addplot [very thick,mark=none,uniaccentblue,samples=2,domain=0:450]{1.308};
\addplot [very thick,mark=none,uniaccentblue,samples=2,domain=0:450]{0};
\legend{BestAlgorithm, RandomForest, BestAlgorithm (mean), RandomForest (mean)}
\end{axis}
\end{tikzpicture}%
~%
%
\begin{tikzpicture}
\begin{axis}[
	title=log scale (base 2),
	height=\textwidth,
	width=\textwidth,
	ylabel={BestThreeLoss}, 
	xlabel={Data Set},
	ymode=log,
	log basis y={2}
]
\addplot [mark=*,only marks,mark size=1pt] table [x=ID,y=BestThreeLoss,col sep=semicolon] {data/BestAlgorithmRanker_metaData_small_allPerformanceValues.csv};
\addplot [mark=*,only marks,mark size=1pt,uniaccentblue] table [x=ID,y=BestThreeLoss,col sep=semicolon] {data/RandomForestRanker_metaData_small_allPerformanceValues.csv};
\addplot [very thick,mark=none,black,samples=2,domain=0:450]{2.440};
\addplot [very thick,mark=none,uniaccentblue,samples=2,domain=0:450]{1.308};
\addplot [very thick,mark=none,uniaccentblue,samples=2,domain=0:450]{0};
\legend{BestAlgorithm, RandomForest, BestAlgorithm (mean), RandomForest (mean)}
\end{axis}
\end{tikzpicture}
}
\caption{The Best Three Loss for the Best Algorithm Baseline and Random Forest Ranker (Full Meta Data)}
\label{fig:lossGraphics}
\end{figure}

>>>>

% How base performance data generated
The performance data of the classifiers for the selected data sets was generated on a linux cluster, each node consisting of two Intel(R) Xeon(R) E5-2670 processors running at 2.6GHz, 16GB RAM with usage of up to 150 nodes. There was a timeout of eight hours set for the computation of a single performance value of a classifier on a data set. For each of the classifiers, the predictive accuracy (the percentage of correctly classified instances in the data set) was recorded on each data set by means of five times stratified Monte-Carlo cross validation with a 70/30 split in training and test data. The stratified split was generated with the help of JAICore, 'A Java library for infrastructural classes and basic algorithms in the area reasoning, search, and machine learning' [Reference Missing]. When a timeout occurred or the evaluation failed otherwise, the predictive accuracy was set to zero.

% How rest of the evaluation carried out
All other experiments where carried out on a Windows Machine running Windows 10 Education (version 1709) with a Intel(R) Core(TM) i5-4200U CPU running at 1.60GHz (2.30GHz max) with 12 GB RAM. It is notable that especially the time-related results are to be viewed in regard to this setup. No timeouts where used in this evaluation, and the rankers where evaluated on the training data by means of leave-one-out estimation, that is for the n data sets for which performance values where recorded, they where trained with the meta data - performance examples for n - 1 data sets, and queried for the remaining data set. As n = 448 in this case, the detailed values for each data set are not included in the discussion of the results. They can, however, be found in the supplementary material, offering clues to how each ranker performed in the prediction of a ranking for each of the used data sets, identified by their ID on OpenML. Furthermore, the evaluation was carried out twice for all measures: once with the full meta data used, and one time using only the meta features not generated by landmarkers, that is the ones that require probing to be carried out for a new data set.

% Implementation of measures
For the computation of the Mann-Whitney U and Kendall rank correlation coefficient, Apache Commons implementations where used. Regarding the Mann-Whitney U, the strategy for ties as set to averaging, and the NaNs where set to be removed before the analysis was started.

@@

As for the rank correlation most of the rankers had a significant advantage over the baseline, it is interesting to compare them among each other. From Table \ref{tab:evaluationResults} it is clear that the regression rankers are superior to the preference rankers, and this advantage is significant, as the weakest of the regression rankers, REPTree, is still significantly better than the strongest label ranking approach, instance based label ranking with Kemeny-Young rank aggregation ($p \approx 4.305E-6$). Out of the four regression rankers, three approaches are furthermore significantly better than REPTree.

Compared to the evaluation on the full set of meta data, the results on the reduced set without probing are less distinctive. For the Kendall Rank correlation, linear regression, M5P and random forest are significantly better than the baseline, and for label ranking, both of the instance based approaches with non-default parameters as well, while the other label ranking approaches are significantly worse than the baseline. In total, random forest is still the best alternative with a correlation of 0.439, slightly worse than the correlation achieved on the full meta data. For both loss measures however, while the disadvantages of the label ranking approaches remain, none of the eight alternatives achieves a significantly better loss or best three loss than the baseline anymore.

With reduced meta data, out of the alternatives that pose a significant advantage compared to the baseline, only random forest is significantly better than all other approaches, the others are not significantly different. With loss and best three loss, no such statement can be made.

Summarizing the observed results, in general, a regression based approach shows an indication to be stronger than a preference based approach. Specifically, using random forest as a regression model is the best solution observed in the evaluation, and label ranking by pairwise comparison the weakest, as it falls behind the other ranking approaches and the baseline often A surprising result is that the label ranking approaches achieve very similar results with or without meta data.

@@

\tikzsetnextfilename{LossGraphics}
\begin{figure}
%\pgfplotsset{width=\textwidth}
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}
\begin{axis}[
	title=regular scale,
	height=\textwidth,
	width=\textwidth,
	ylabel={BestThreeLoss}, 
	xlabel={Data Set},
]
\addplot [mark=*,only marks,mark size=1pt] table [x=ID,y=BestThreeLoss,col sep=semicolon] {data/BestAlgorithmRanker_metaData_small_allPerformanceValues.csv};
\addplot [mark=*,only marks,mark size=1pt,uniaccentblue] table [x=ID,y=BestThreeLoss,col sep=semicolon] {data/RandomForestRanker_metaData_small_allPerformanceValues.csv};
\addplot [very thick,mark=none,black,samples=2,domain=0:450]{2.440};
\addplot [very thick,mark=none,uniaccentblue,samples=2,domain=0:450]{1.308};
\addplot [very thick,mark=none,uniaccentblue,samples=2,domain=0:450]{0};
\legend{BestAlgorithm, RandomForest, BestAlgorithm (mean), RandomForest (mean)}
\end{axis}
\end{tikzpicture}%
~%
%
\begin{tikzpicture}
\begin{axis}[
	title=log scale (base 2),
	height=\textwidth,
	width=\textwidth,
	ylabel={BestThreeLoss}, 
	xlabel={Data Set},
	ymode=log,
	log basis y={2}
]
\addplot [mark=*,only marks,mark size=1pt] table [x=ID,y=BestThreeLoss,col sep=semicolon] {data/BestAlgorithmRanker_metaData_small_allPerformanceValues.csv};
\addplot [mark=*,only marks,mark size=1pt,uniaccentblue] table [x=ID,y=BestThreeLoss,col sep=semicolon] {data/RandomForestRanker_metaData_small_allPerformanceValues.csv};
\addplot [very thick,mark=none,black,samples=2,domain=0:450]{2.440};
\addplot [very thick,mark=none,uniaccentblue,samples=2,domain=0:450]{1.308};
\addplot [very thick,mark=none,uniaccentblue,samples=2,domain=0:450]{0};
\legend{BestAlgorithm, RandomForest, BestAlgorithm (mean), RandomForest (mean)}
\end{axis}
\end{tikzpicture}
}
\caption{The best three loss for the baseline compared to the random forest ranker (full meta data).}
\label{fig:lossGraphics}
\end{figure}

### 2018-02-14 Writing/Thesis/content/chapter-fundamentals.tex
First draft

\subsubsection{Ranking and Ordering}
To clarify the meaning of ranking and ordering, we return to the gardening example. After spending some time studying the different flowers, the gardener has realized he prefers certain iris species to others.

Ranking = Permutation of the actual labels!
%TODO {fill}
Ordering =

A ranking for a set of items defines a strict total order

% Gardener example ranking his favourite plants
The ordering sorts the labels according to their given score. It thus may also implicitly be given by ordering the items themselves, as the labels are aliases for them. Therefore, an ordering for a set of labels $\lbrace 1\dots k\rbrace$ is a permutation $\pi$ of the labels where $\pi(i)$ is the

It must be noted that the definitions given here are not transferable to the general case of a ranking or ordering problems, since a strict total order is not required in all contexts for either ranking or ordering.

@@

The values in table \ref{tab:ranking_vs_ordering} result in the ranking $[2,3,1]$, whereas the ordering is $[3,1,2]$.

\subsubsection{Label Ranking}
Similar to classification, in label ranking there is a finite set of labels $Y=\lbrace y_1,\dots,y_n \rbrace$, but it is not the output space. Instead, \citeauthor{DBLP:books/daglib/0025729} define the target function for label ranking as $f:X\rightarrow S_Y$ where the output space $S_Y$ contains all permutations over the set of labels $Y$ \cite{DBLP:books/daglib/0025729}. An instance $x \in X$ therefore is assigned a permutation $y_{\pi_x^i}$

Although called label ranking actually label ordering!

original algorithms selection problem \cite{rice1976algorithm} can be viewed as a learning problem itself

\subsection{Meta Learning}
% Context is meta learning -> what does that mean and what are the consequence
Meta leThe previous discussion of machine learning problems has conceptually been about the fitting of different machine learning algorithms to problems in the form of data sets. In the context of meta learning, is 'possible to learn about the learning process
itself, and in particular that a system could learn to profit from previous
experience to generate additional knowledge that can simplify the automatic
selection of efficient models summarizing the data' \cite{brazdil2008metalearning}.

% Algorithm selection problem

@@

general measure predictive accuracy (e.g. go back to example gardener model is not perfect so miss classifies some instancs if want to know how good exactly, how find out?

- estimation using cross validation (monte carloMCCV)

- estimation using leave one out

@@

%TODO include example ranking

@@

%TODO include example loss calculatio

\subsection{Mann-Whitney U Test}
%TODO describe & don't forget example
The measures previously presented can indicate if one alternative is superior to another regarding that measure, but one cannot conclude from those results alone whether this advantage is significant. The question is whether the null-hypothesis that two distributions \cite{mann1947test}

>>>>

\subsubsection{Ranking and Ordering}
To clarify the meaning of ranking and ordering, we return to the gardening example. After spending some time studying the different flowers, the gardener has realized he prefers certain iris species to others. He expresses his preference for a plant in a score, and labels all the plants in no particular order (see Fig. \ref{tab:ranking_vs_ordering}). An \textit{ordering} then is the result of simply ordering the labels in decreasing order of their respective score. It thus can also be given implicitly by just ordering the original classes, the flower names, as such being a permuation of the labels. A ranking, on the other hand, assigns a fixed position to each label; for instance the rank of 'iris virginica' will always be in the first position in the ranking, that is the number in that position denotes the \textit{rank} of the label, which is in turn defined as the label's position in the ordering. Therefore the values in the example in Table \ref{tab:ranking_vs_ordering} result in the ranking $[2,3,1]$, whereas the ordering is $[3,1,2]$. It is notable that these can be the same, and are easily convertible. However, in the remainder of this thesis mostly the term ranking is used to identify both ranking and ordering interchangeably, especially as usually only the term 'ranking' is used.

@@

\subsubsection{Label Ranking}
Similar to classification, in label ranking there is a finite set of labels $Y=\lbrace y_1,\dots,y_n \rbrace$, but it is not the output space. Instead, \citeauthor{DBLP:books/daglib/0025729} define the target function for label ranking as $f:X\rightarrow S_Y$ where the output space $S_Y$ contains all permutations over the set of labels $Y$ \cite{DBLP:books/daglib/0025729}. To come back to the topic of rankings and orderings, it becomes clear now that even though it is called label ranking, the labels are actually ordered. An example of label ranking could be, to stay in the domain of gardening, to learn the preferences of different gardeners towards iris plants. A label ranking data set used to learn this connection could, for example, map the favourite colour and height of a gardener to a ranking of the three iris plants.

\subsection{Meta Learning}
% Context is meta learning -> what does that mean and what are the consequence
The previous discussion of machine learning problems has conceptually been about the fitting of different machine learning algorithms to problems in the form of data sets. In the context of meta learning regarding machine learning, it is researched whether it is 'possible to learn about the learning process itself, and in particular that a system could learn to profit from previous experience to generate additional knowledge that can simplify the automatic selection of efficient models summarizing the data' \cite{brazdil2008metalearning}. That is, the original algorithm selection problem \cite{rice1976algorithm} is viewed as a learning problem itself that is to be tackled.

The approach of this thesis clearly falls under that category, as it is trying to learn whether a connection exists beween properties of a data set, that is its meta features, and the performance values of certain classification algorithms, and how well it can be exploited by regression and preference models respetively, if it exists. The benefit from meta learning in the conext of machine learning is that learning how properties of data sets influence the learning process of the classifiers could save much work; Often, much work is put into the fitting of an algorithm to a specific data set, but it is not transferred to other poblems that may be of the same nature.

@@

To find out how well a classifier performs, it first needs to be decided which measure consider. There are a few possibilities for this, however in this thesis the term performance measure refers to the predictive accuracy of a classifier, that is the percentage of classes that a classifier correctly identified in a given data set. But since using the whole data set to train a classifier, that is build a model, would skew the analysis - the classifier has already 'seen' these examples - a method for dividing data sets into train and test sets is needed.

A popular approach is using a stratified split, for example splitting the data into 70\% test and 30\% training data, with the important aspect that both sub sets have equal (or at least very similar) ratios of classes [Reference Missing]. The learning algorithm is then trained with one part of the data (usually the larger), and its performance tested on the other. To get more stable results, often the process is repeated: The data is divided in k folds, k times, the learning algorithm is trained on k-1 folds, and tested on the left out subset. Results are averaged. This process is then called k-fold (stratified) cross validation [Reference Missing]. A variant of this is Monte Carlo crossvalidation, where after each time the algorithm has been trained and evaluated, a new split is computed [Reference Missing]. This especially causes the fact that the number of folds does not dictate the ratio of the split; one could for example have 5-folds Mone Carlo crossvalidation with a 70\%/30\% train/test split.

An alternative to using a stratified split is using a method called leave-one-out. This method essentially is, for a data set with n instances, n-fold crossvalidation (not stratified, naturally). This means that the algorithm is trained on n-1 instances and evaluated on the instance that has been left out.

@@

@@

\subsection{Mann-Whitney U Test}
The measures previously presented can indicate if one of the ranking alternatives is superior to another regarding that measure, but one cannot conclude from those results alone whether this advantage is significant. The question here is whether the null-hypothesis that two distributions belong to the same population can be rejected. This can be determined with the help of the Mann-Whitney U, a non-parametric hypthesis test \cite{mann1947test}. The Mann-Whitney U is computed by, for two distributions one wants to analyze, counting how many times each value frome one distribution is lower than another value from the other distribution. If the U is very high (or low, depending on the hypothesis), the null hypothesis can be rejected. Whether this significance level is reached is expressed by the p-value: for a p-value below 0.05, The null hypothesis can be rejected. Therefore, if two distributions, for example regarding the loss values of two ranking approaches, suggest that one is superior to the other, and the p-value for the Mann-Whitney U is <= 0.05, this suggests that the difference is statistically significant.
